
---
title: "Household catch species diversity vs CFR Biomonitoring species diversity"
output: 
  pdf_document:
    latex_engine: xelatex
mainfont: Arial

knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/github/cambodia_biodiversity/output") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```



```{r packages, echo = FALSE, include = FALSE}
library(Hmisc) # install.packages("Hmisc") # this package does makes summarize not recognize group_by(), use summarise with an s instead.
library(lubridate)
library(ggalluvial) # install.packages("ggalluvial")
library(alluvial) #install.packages("alluvial")
library(tidyverse)
library(viridis) #install.packages("viridis")
library (hrbrthemes) #install.packages("hrbrthemes")

projectFolder  <- "~/github/cambodia_biodiversity"
data <- file.path(projectFolder, "data/processed/")
output <-file.path(projectFolder, "output/")

```


```{r read_files}
# bring in csv files generated in stata

# catch data
c <- read.csv(file.path(data, "ccm_q7.csv"), na.strings = c("", "NA")) 
c <- as_tibble(c)

# biomonitoring data
b <- read.csv(file.path(data, "biom_formb.csv"), na.strings = c("", "NA"))
b <- as_tibble(b)

# cfr types
cfrtype <- read.csv(file.path(data, "cfr_category.csv"), na.strings = c("", "NA"))
cfrtype <- as_tibble(cfrtype)

# all dates
dates <- as_tibble(read.csv(file.path(data, "dates.csv"), na.strings = c("", "NA")))

```


### Biomonitoring X Catch Plots

Below are plots of household catch species diversity (count of species) against CFR level species diversity (count of species). 

The latter two figures aggregate by 6 month time blocks in slightly different ways. 

Below each figure is the Pearson correlation coefficient and pvalue. Pvalues are tinyyyyy.  
\
\


```{r}
# Create a HOUSEHOLD-level data frame that contains the number of species caught by households associated with that CFR

cfr_catch <- c %>% 
  select(cfrid, hhid, scode_ccm) %>% 
  distinct(cfrid, hhid, scode_ccm) %>% 
  group_by(cfrid, hhid) %>% 
  summarise(catch_species = n())


# Create a CFR-level data frame that cnotains the number of species caught during CFR biomonitoring
cfr_biom <- b %>% 
  select(cfrid, scode_biom) %>% 
  distinct(cfrid, scode_biom) %>% # dropping duplicates in terms of CFR and species
  group_by(cfrid) %>% 
  summarise(biom_species = n())


# Combine data frames
cfr_biodiv <- full_join(cfr_biom, cfr_catch, by = "cfrid") 

# Check correlation stats
corr <- cfr_biodiv %>% 
  select(biom_species, catch_species) %>% 
  as.matrix() %>% 
  Hmisc::rcorr(type = "spearman") %>% 
  broom::tidy() %>% 
  knitr::kable()


# Plot
cfr_biodiv %>% 
  ggplot(aes(x = biom_species, y = catch_species)) +
  geom_point() +
  geom_smooth(method = "lm")+
  ggtitle("Catch at HH level; Biomonitoring at CFR level")

corr
```


```{r}

## AGGREGATE BY 6 MONTH TIME BLOCK (CATCH ONLY)

# Merge all month/year combos into household level CFR data frame of catch data
alldates_catch <- c %>% 
  select(hhid, cfrid, year, month, scode_ccm) %>% 
  full_join(dates, by = c("year", "month")) %>% 
  arrange(year, month)
  # Note that hhid, cfrid and scode are all NA for months where catch data was not collected

# Merge all month/year combos into CFR-level data frame of biomonitoring data
alldates_biom <- b %>% 
  select(cfrid, year, month, scode_biom) %>% 
  full_join(dates, by = c("year", "month")) %>% 
  arrange(year, month)

# Create a HOUSEHOLD-level data frame that contains the number of species caught by households associated with that CFR--AGGREGATED BY 6 MONTH TIME BLOCK
cfr_catch6 <- alldates_catch %>% 
  select(cfrid, hhid, scode_ccm, timeblock6) %>% 
  distinct(cfrid, hhid, scode_ccm, timeblock6) %>% 
  group_by(cfrid, hhid, timeblock6) %>% 
  summarise(catch_species = n()) %>% 
  arrange(timeblock6) # NA values in timeblock 6 are Sept/Oct/Nov 2015

# Create a CFR-level data frame that contains the number of species caught during CFR biomonitoring--AGGREGATED BY 6 MONTH TIME BLOCK
cfr_biom <- b %>% 
  select(cfrid, scode_biom) %>% 
  distinct(cfrid, scode_biom) %>% # dropping duplicates in terms of CFR and species
  group_by(cfrid) %>% 
  summarise(biom_species = n())

# Combine data frames
cfr_biodiv6 <- full_join(cfr_biom, cfr_catch6, by = "cfrid") %>% 
  drop_na(cfrid, timeblock6)

# Check correlation stats
corr <- cfr_biodiv6 %>% 
  ungroup() %>% 
  select(biom_species, catch_species) %>% 
  as.matrix() %>% 
  Hmisc::rcorr(type = "spearman") %>% 
  broom::tidy() %>% 
  knitr::kable()


# Plot
cfr_biodiv6 %>% 
  ggplot(aes(x = biom_species, y = catch_species)) +
  geom_point() +
  geom_smooth(method = "lm")+
  ggtitle("Catch at HH-timeblock6 level; Biomonitoring at CFR level")

corr
```


```{r}

## AGGREGATE BY 6 MONTH TIME BLOCK (CATCH AND BIOMONITORING)

# Merge all month/year combos into household level CFR data frame of catch data
alldates_catch <- c %>% 
  select(hhid, cfrid, year, month, scode_ccm) %>% 
  full_join(dates, by = c("year", "month")) %>% 
  arrange(year, month)
  # Note that hhid, cfrid and scode are all NA for months where catch data was not collected

# Merge all month/year combos into CFR-level data frame of biomonitoring data
alldates_biom <- b %>% 
  select(cfrid, year, month, scode_biom) %>% 
  full_join(dates, by = c("year", "month")) %>% 
  arrange(year, month)

# Create a HOUSEHOLD-level data frame that contains the number of species caught by households associated with that CFR--AGGREGATED BY 6 MONTH TIME BLOCK
cfr_catch6 <- alldates_catch %>% 
  select(cfrid, hhid, scode_ccm, timeblock6) %>% 
  distinct(cfrid, hhid, scode_ccm, timeblock6) %>% 
  group_by(cfrid, hhid, timeblock6) %>% 
  summarise(catch_species = n()) %>% 
  arrange(timeblock6) # NA values in timeblock 6 are Sept/Oct/Nov 2015

# Create a CFR-level data frame that contains the number of species caught during CFR biomonitoring--AGGREGATED BY 6 MONTH TIME BLOCK
cfr_biom6 <- alldates_biom %>% 
  select(cfrid, scode_biom, timeblock6) %>% 
  distinct(cfrid, scode_biom, timeblock6) %>% # dropping duplicates in terms of CFR and species
  group_by(cfrid, timeblock6) %>% 
  summarise(biom_species = n())

# Combine data frames
cfr_biodiv6 <- full_join(cfr_biom6, cfr_catch6, by = c("cfrid", "timeblock6")) %>% 
  drop_na(cfrid, timeblock6)

# Check correlation stats
corr <- cfr_biodiv6 %>% 
  ungroup() %>% 
  select(biom_species, catch_species) %>% 
  drop_na() %>% 
  as.matrix() %>% 
  Hmisc::rcorr(type = "spearman") %>% 
  broom::tidy() %>% 
  knitr::kable()


# Plot
cfr_biodiv6 %>% 
  ggplot(aes(x = biom_species, y = catch_species)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ggtitle("Catch at HH-timeblock6 level; Biomonitoring at CFR-timeblock6 level")

corr 
```

### Biomonitoring X Consumption 

For all: Consumption at HH level; Biomonitoring at CFR level. Ignoring time dimension. 

Biomonitoring presence/absence X Consumption presence/absence
```{r}

# Generate CFR-level biomonitoring file
cfr_biom <- b %>% 
  distinct(cfrid, scode_biom) %>% 
  mutate(scode = scode_biom) %>% 
  group_by(cfrid) %>% 
  summarise(biom_species = n())

# Generate household-level consumption file
hh_consumption <- c %>% 
  distinct(hhid, cfrid,  scode_ccm, atefresh) %>% 
  mutate(scode = scode_ccm) %>% 
  group_by(cfrid, hhid) %>% 
  summarise(consumption_species = n())
  
# Combine files
biom_consumption <- full_join(cfr_biom, hh_consumption, by = c("cfrid"))

# Check correlation stats
corr <- biom_consumption %>% 
  select(biom_species, consumption_species) %>% 
  as.matrix() %>% 
  Hmisc::rcorr(type = "spearman") %>% 
  broom::tidy() %>% 
  knitr::kable()

# Plot
plot <- biom_consumption %>% 
  ggplot(aes(x = biom_species, y = consumption_species)) +
  geom_point() +
  geom_smooth(method = "lm")+
  ggtitle("Consumption at HH level; Biomonitoring at CFR level")

plot
corr


```






### Catch X Consumption

