
---
title: "Cambodia biodiversity paper notes -- Log rank abundance plots--August 8, 2022"
output: 
  pdf_document:
    latex_engine: xelatex
mainfont: Arial

knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/github/cambodia_biodiversity/output") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```



```{r packages, echo = FALSE, include = FALSE}
library(Hmisc) # install.packages("Hmisc") # this package makes summarize not recognize group_by(), use summarise with an s instead. Sometimes ungroup() is needed when you least expect it.
library(lubridate)
library(ggalluvial) # install.packages("ggalluvial")
library(alluvial) #install.packages("alluvial")
library(tidyverse)
library(viridis) #install.packages("viridis")
library (hrbrthemes) #install.packages("hrbrthemes")
library(kableExtra)

library(vegan) # install.packages("vegan")
library(BiodiversityR) #install.packages("BiodiversityR")

library(pheatmap) #install.packages("pheatmap")

projectFolder  <- "~/github/cambodia_biodiversity"
data <- file.path(projectFolder, "data/processed/")
output <-file.path(projectFolder, "output/")

```


```{r read_files}
# bring in csv files generated in stata

# cfr types
cfrtype <- as_tibble(read.csv(file.path(data, "cfr_category.csv"), na.strings = c("", "NA")))

# all dates
dates <- as_tibble(read.csv(file.path(data, "dates.csv"), na.strings = c("", "NA")))

# species codes
scode <- as_tibble(read.csv(file.path(data, "species_codes.csv"), na.strings = c("", "NA")))

# catch data
c <- as_tibble(read.csv(file.path(data, "ccm_q7.csv"), na.strings = c("", "NA")))

# biomonitoring data
b <- as_tibble(read.csv(file.path(data, "biom_formb.csv"), na.strings = c("", "NA")))

# traits data
t <- as_tibble(read.csv(file.path(data, "traits.csv"), na.strings = c("", "NA")))
  # merge codes into 


```

``` {r, eval = FALSE}

Questions we might answer: 

How many species make it from biom to catch? 
Of species that make it through, what share of biomass makes it through? 
How is the distribution of abundance different from biom to catch? Rank abundance curve. https://en.wikipedia.org/wiki/Rank_abundance_curve

How many species make it from catch to plate?
Of species that make it from catch to plate, what share of biomass makes it through? Remember that consumption biomass is biologically limited. 
How is the distribution of abundance different from catch to plate?
```

```{r}

## Create a data frame that contains rank abundance data by biom, catch, cons

# biom
bsum <- b %>% 
  select(sname_biom, totalweight_biom) %>% 
  group_by(sname_biom) %>% 
  summarise(weight = sum(totalweight_biom)) %>% 
  # adding rank
  arrange(-weight) %>%
  mutate(rank = seq(1,length(weight)),
         type = "Biomonitoring",
         sname = sname_biom)

# catch
csum <- c %>% 
  select(sname_ccm, catch_iweight) %>% 
  group_by(sname_ccm) %>% 
  summarise(weight = sum(catch_iweight)) %>% 
  arrange(-weight) %>% 
  mutate(rank = seq(1,length(weight)),
         type = "Catch",
         sname = sname_ccm)

# consumption
asum <- c %>% 
  select(sname_ccm, atefresh_iweight) %>% 
  group_by(sname_ccm) %>% 
  summarise(weight = sum(atefresh_iweight)) %>% 
  arrange(-weight) %>% 
  mutate(rank = seq(1,length(weight)),
         type = "Consumption",
         sname = sname_ccm)

ra <- bsum %>% 
  bind_rows(csum) %>% 
  bind_rows(asum) %>% 
  filter(rank < 100) %>% 
  mutate(weight = log1p(weight))
  


```

``` {r}
# Make Rank Abundance plots
# I experimented with the rankabundance function and it seemed to be more trouble than it was worth, so making them by hand

# Let's make the plot in ggplot where I have more control
# define theme
simple <- theme(
  panel.background = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.line = element_line("gray25"),
  #text = element_text(size = 12, family="Arial"),
  axis.text = element_text(size = 10, colour = "gray25"),
  axis.title = element_text(size = 14, colour = "gray25"),
  legend.title = element_text(size = 14),
  legend.text = element_text(size = 14),
  legend.key = element_blank()) +
  labs(y = "log1p(weight)", x = "Ranked Species")



# Catch
ra_catch <- ra %>% 
  filter(type == "Catch") %>% 
  ggplot(aes(x = rank, y = weight)) + 
  geom_line() +
  geom_point(size=2, alpha=0.3, color = "#009999") +
  geom_text(aes(label = ifelse(rank <= 5, sname,'')),
            hjust = 0, vjust = 0,
            size = 3) +
  geom_text(aes(label = ifelse(rank == 6, sname,'')),
            hjust = 0, vjust = -0.5,
            size = 3) +
  geom_text(aes(label = ifelse(rank == 7, sname,'')),
            hjust = 0, vjust = 0.5,
            size = 3) +
  simple +
  ggtitle("Log Rank Abundance--Catch")

# Consumption
ra_cons <- ra %>% 
  filter(type == "Consumption") %>% 
  ggplot(aes(x = rank, y = weight)) + 
  geom_line() +
  geom_point(size=2, alpha=0.3, color = "#009999") +
  geom_text(aes(label = ifelse(rank <= 4, sname,'')),
            hjust = 0, vjust = 0,
            size = 3) +
  geom_text(aes(label = ifelse(rank == 5 | rank == 7, sname,'')),
            hjust = 0, vjust = 0.5,
            size = 3) +
  geom_text(aes(label = ifelse(rank == 6, sname,'')),
            hjust = 0, vjust = -0.5,
            size = 3) +
  simple +
  ggtitle("Log Rank Abundance--Consumption")
  
```

I calculated rank abundance at the system level (biomonitoring data), catch level and consumption level. 

I labeled the top 7 species in each plot.

The scales are quite different, making it difficult to put them side-by-side. 

I plotted only the top 70 species. This reduced the very long right tail in all plots. 

The shape of the distributions are all quite similar to my eyes.

The most caught and consumed species (Channa striata) is also the most common in biomonitoring. The second most commonly caught and consumed species (Anambas testudineus) does not appear in the top 7 of biomonitoring. It could still be a very common species, but this indicates that there's some targeting going on, as expected.

The second, third and fourth most abundant species in the biomonitoring data do not appear in the top 7 of catch and consumption. 



\
\
\
\

``` {r}
#ra_biom
ra_catch
ra_cons
```



```{r, eval = FALSE}

# Let's try making a heat map using pheatmap

# Starting with instructions from: https://environmentalcomputing.net/graphics/multivariate-vis/heatmaps/

# Need to create a dataset where each row is a species and each column is the abundance of that species per CFR

cfr_weights <- as.data.frame(b) %>% 
  group_by(cfrid, sname_biom) %>% 
  summarise(weight = sum(totalweight_biom)) %>% 
  pivot_wider(names_from = cfrid, values_from = weight) %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  column_to_rownames(var = "sname_biom")

#regmap <- pheatmap(cfr_weights)

cfr_weights10 <- log1p(cfr_weights)
#pheatmap(cfr_weights10)

logmap <- pheatmap(cfr_weights10, method = c("pearson"), 
         #cellwidth = 5, 
         #cellheight = 3, 
         fontsize = 2.5, 
         width = 100, height = 400) # doesn't seem to change anything

#logmap
```


